### 2016-09-08 analysis-note-v2/Appendix.tex
update the answers and minor text cleanings in the note
Clarification on calibration and clean up

~~~~~The exclusive selection of the experimental $e^{4}He\pi^{0}$ and $ep\pi^{0}$ events require the detection of only one good electron, one good $\pi^{0}$ in the topology ICIC or ICEC, and one good $^{4}He$ track in the coherent channel or one good proton in the incoherent channel case. Furthermore, in order to ensure that this is a deep process we apply a set of initial requirements. The exclusivity of the reaction is ensured by a set of exclusivity cuts like for the DVCS channels. These requirements and exclusivity cuts are presented for the case of the coherent $e^{4}He\pi^{0}$ events are:

\paragraph{Initial criteria}
~\\
~\\
These requirements are made to ensure that the selected events occurre at the partonic level:
\begin{itemize}
\item High virtuality of the exchanged photon ($Q^{2}$>1 $GeV^{2}$).
\item High energy of the emitted $\pi^{0}$ ($E_{\pi^{0}}$ > 2 GeV).
\item The invariant mass of the virtual photon and the target proton is greater than 2 GeV$^{2}$/c$^{2}$ in order to avoid the baryons resonances region.
\item The transfer momentum squared ($-t$) between the initial target and the recoil one is greater than the minimum allowed one ($t_{min}$) defined by the kinematics of the incoming and the scattered electrons. The definition of $t_{min}$ for each channel can be found in the corresponding DVCS channel selection presented previously.
\end{itemize}

\paragraph{Exclusivity requirements}
~\\
~\\The exclusivity of the selected $e^{4}He\pi^{0}$ events is done with the following cuts:
\begin{itemize}
\item The coplanarity cut ($\Delta \phi$) between the recoil $^{4}He$ and the produced $\pi^{0}$.
\item The missing energy, mass and transverse momentum cuts in the configuration $e^{4}He\pi^{0}X$.
\item The missing mass cut in the configuration $e^{4}HeX$.
\item The missing mass cut in the configuration $e\pi^{0}X$.
\item The cone angle cut between $e^{4}HeX$ and the reconstructed $\pi^{0}$.
\end{itemize}
  The same procedure holds for the case of the incoherent $ep\pi^{0}$ events. In the following two subsections, the results of the two channels selection are presented.
~\newpage
\section{$e^{4}He\pi^{0}$ exclusivity cuts}
~~~~The events which pass the following exclusivity cuts are assumed to be good $e^{4}He\pi^{0}$ events.

@@

\chapter{ALU tables}
\begin{table}[!h]
   \begin{center}
      \begin{tabular}{||l|l|l|l|l||}
         \hline
 $<Q^{2}>$ & $<x_{B}>$ & $<-t>$ & $<\phi>$ & $A_{LU}$ $\pm$ stat. $\pm$ syst.\\
         \hline
  1.143 & 0.136 & 0.096 &  23.4  &  0.1407   $\pm$ 0.0800   $\pm$  0.0257 \\
  1.143 & 0.136 & 0.096 &  59.9  &  0.3563   $\pm$ 0.0687   $\pm$  0.0188 \\
  1.143 & 0.136 & 0.096 &  97.6  &  0.4241   $\pm$ 0.0870   $\pm$  0.0403 \\
  1.143 & 0.136 & 0.096 &  139.9 &  0.1576   $\pm$ 0.1344   $\pm$  0.0267 \\
  1.143 & 0.136 & 0.096 &  179.1 &  0.1496   $\pm$ 0.1390   $\pm$  0.0278 \\
  1.143 & 0.136 & 0.096 &  218.9 & -0.1361   $\pm$ 0.1123   $\pm$  0.0248 \\
  1.143 & 0.136 & 0.096 &  263.5 & -0.3384   $\pm$ 0.0922   $\pm$  0.0389 \\
  1.143 & 0.136 & 0.096 &  302.9 & -0.1731   $\pm$ 0.0722   $\pm$  0.0263 \\
  1.143 & 0.136 & 0.096 &  337.5 & -0.3439   $\pm$ 0.0744   $\pm$  0.0343 \\
  \hline                                                           
  1.423 & 0.172 & 0.099 &  19.9  &  0.1538   $\pm$ 0.0650   $\pm$  0.0267 \\
  1.423 & 0.172 & 0.099 &  57.3  &  0.2941   $\pm$ 0.0684   $\pm$  0.0250 \\
  1.423 & 0.172 & 0.099 &  95.4  &  0.4352   $\pm$ 0.0997   $\pm$  0.0426 \\
  1.423 & 0.172 & 0.099 &  138.3 &  0.1251   $\pm$ 0.1418   $\pm$  0.0250 \\
  1.423 & 0.172 & 0.099 &  181.2 &  0.2139   $\pm$ 0.1813   $\pm$  0.0302 \\
  1.423 & 0.172 & 0.099 &  220.3 & -0.2494   $\pm$ 0.1646   $\pm$  0.0344 \\
  1.423 & 0.172 & 0.099 &  263.4 & -0.3187   $\pm$ 0.1128   $\pm$  0.0366 \\
  1.423 & 0.172 & 0.099 &  302.5 & -0.3429   $\pm$ 0.0695   $\pm$  0.0383 \\
  1.423 & 0.172 & 0.099 &  338.1 & -0.1382   $\pm$ 0.0680   $\pm$  0.0261 \\
  \hline                                                                    
  1.902 & 0.224 & 0.107 &  21.0  &  0.0904   $\pm$ 0.0630   $\pm$  0.0234 \\
  1.902 & 0.224 & 0.107 &  56.6  &  0.3647   $\pm$ 0.0654   $\pm$  0.0185 \\
  1.902 & 0.224 & 0.107 &  95.7  &  0.1701   $\pm$ 0.1072   $\pm$  0.0173 \\
  1.902 & 0.224 & 0.107 &  139.6 &  0.0994   $\pm$ 0.1732   $\pm$  0.0170 \\
  1.902 & 0.224 & 0.107 &  179.5 & -0.0855   $\pm$ 0.2165   $\pm$  0.0126 \\
  1.902 & 0.224 & 0.107 &  221.6 & -0.1338   $\pm$ 0.1825   $\pm$  0.0152 \\
  1.902 & 0.224 & 0.107 &  262.7 & -0.2772   $\pm$ 0.1185   $\pm$  0.0260 \\
  1.902 & 0.224 & 0.107 &  303.5 & -0.2306   $\pm$ 0.0745   $\pm$  0.0207 \\
  1.902 & 0.224 & 0.107 &  339.2 & -0.1958   $\pm$ 0.0607   $\pm$  0.0163 \\
         \hline \hline
      \end{tabular}
      \caption{ The coherent $A_{LU}$ in $Q^2$ bins}
      \label{table:Coh_Q2_BSA}
   \end{center}
\end{table}

>>>>

The exclusive selection of the experimental $e^{4}He\pi^{0}$ and $ep\pi^{0}$ events require the detection of only one good electron, one good $\pi^{0}$ in the topology ICIC or ICEC, and one good $^{4}He$ track in the coherent channel or one good proton in the incoherent channel case. Furthermore, in order to ensure that this is a deep process we apply a set of initial requirements. The exclusivity of the reaction is ensured by a set of exclusivity cuts like for the DVCS channels. These requirements and exclusivity cuts are presented for the case of the coherent $e^{4}He\pi^{0}$ events are:

\paragraph{Initial criteria} ~\\
These requirements are made to ensure that the selected events occurre at the partonic level:
\begin{itemize}
\item High virtuality of the exchanged photon ($Q^{2}$>1 $GeV^{2}$).
\item High energy of the emitted $\pi^{0}$ ($E_{\pi^{0}}$ > 2 GeV).
\item The invariant mass of the virtual photon and the target proton is greater than 2 GeV$^{2}$/c$^{2}$ in order to avoid the baryons resonances region.
\item The transfer momentum squared ($-t$) between the initial target and the recoil one is greater than the minimum allowed one ($t_{min}$) defined by the kinematics of the incoming and the scattered electrons. The definition of $t_{min}$ for each channel can be found in the corresponding DVCS channel selection presented previously.
\end{itemize}

\paragraph{Exclusivity requirements}
~\\
~\\The exclusivity of the selected $e^{4}He\pi^{0}$ events is done with the following cuts:
\begin{itemize}
\item The coplanarity cut ($\Delta \phi$) between the recoil $^{4}He$ and the produced $\pi^{0}$.
\item The missing energy, mass and transverse momentum cuts in the configuration $e^{4}He\pi^{0}X$.
\item The missing mass cut in the configuration $e^{4}HeX$.
\item The missing mass cut in the configuration $e\pi^{0}X$.
\item The cone angle cut between $e^{4}HeX$ and the reconstructed $\pi^{0}$.
\end{itemize}
  The same procedure holds for the case of the incoherent $ep\pi^{0}$ events. In the following two subsections, the results of the two channels selection are presented.
~\newpage
\section{$e^{4}He\pi^{0}$ exclusivity cuts}
The events which pass the following exclusivity cuts are assumed to be good $e^{4}He\pi^{0}$ events.

@@

\chapter{$A_{LU}$ tables}
\begin{table}[!h]
   \begin{center}
      \begin{tabular}{||l|l|l|l|l||}
         \hline
 $<Q^{2}>$ & $<x_{B}>$ & $<-t>$ & $<\phi>$ & $A_{LU}$ $\pm$ stat. $\pm$ syst.\\
         \hline
  1.143 & 0.136 & 0.096 &  23.4  &  0.1407   $\pm$ 0.0800   $\pm$  0.0257 \\
  1.143 & 0.136 & 0.096 &  59.9  &  0.3563   $\pm$ 0.0687   $\pm$  0.0188 \\
  1.143 & 0.136 & 0.096 &  97.6  &  0.4241   $\pm$ 0.0870   $\pm$  0.0403 \\
  1.143 & 0.136 & 0.096 &  139.9 &  0.1576   $\pm$ 0.1344   $\pm$  0.0267 \\
  1.143 & 0.136 & 0.096 &  179.1 &  0.1496   $\pm$ 0.1390   $\pm$  0.0278 \\
  1.143 & 0.136 & 0.096 &  218.9 & -0.1361   $\pm$ 0.1123   $\pm$  0.0248 \\
  1.143 & 0.136 & 0.096 &  263.5 & -0.3384   $\pm$ 0.0922   $\pm$  0.0389 \\
  1.143 & 0.136 & 0.096 &  302.9 & -0.1731   $\pm$ 0.0722   $\pm$  0.0263 \\
  1.143 & 0.136 & 0.096 &  337.5 & -0.3439   $\pm$ 0.0744   $\pm$  0.0343 \\
  \hline                                                           
  1.423 & 0.172 & 0.099 &  19.9  &  0.1538   $\pm$ 0.0650   $\pm$  0.0267 \\
  1.423 & 0.172 & 0.099 &  57.3  &  0.2941   $\pm$ 0.0684   $\pm$  0.0250 \\
  1.423 & 0.172 & 0.099 &  95.4  &  0.4352   $\pm$ 0.0997   $\pm$  0.0426 \\
  1.423 & 0.172 & 0.099 &  138.3 &  0.1251   $\pm$ 0.1418   $\pm$  0.0250 \\
  1.423 & 0.172 & 0.099 &  181.2 &  0.2139   $\pm$ 0.1813   $\pm$  0.0302 \\
  1.423 & 0.172 & 0.099 &  220.3 & -0.2494   $\pm$ 0.1646   $\pm$  0.0344 \\
  1.423 & 0.172 & 0.099 &  263.4 & -0.3187   $\pm$ 0.1128   $\pm$  0.0366 \\
  1.423 & 0.172 & 0.099 &  302.5 & -0.3429   $\pm$ 0.0695   $\pm$  0.0383 \\
  1.423 & 0.172 & 0.099 &  338.1 & -0.1382   $\pm$ 0.0680   $\pm$  0.0261 \\
  \hline                                                                    
  1.902 & 0.224 & 0.107 &  21.0  &  0.0904   $\pm$ 0.0630   $\pm$  0.0234 \\
  1.902 & 0.224 & 0.107 &  56.6  &  0.3647   $\pm$ 0.0654   $\pm$  0.0185 \\
  1.902 & 0.224 & 0.107 &  95.7  &  0.1701   $\pm$ 0.1072   $\pm$  0.0173 \\
  1.902 & 0.224 & 0.107 &  139.6 &  0.0994   $\pm$ 0.1732   $\pm$  0.0170 \\
  1.902 & 0.224 & 0.107 &  179.5 & -0.0855   $\pm$ 0.2165   $\pm$  0.0126 \\
  1.902 & 0.224 & 0.107 &  221.6 & -0.1338   $\pm$ 0.1825   $\pm$  0.0152 \\
  1.902 & 0.224 & 0.107 &  262.7 & -0.2772   $\pm$ 0.1185   $\pm$  0.0260 \\
  1.902 & 0.224 & 0.107 &  303.5 & -0.2306   $\pm$ 0.0745   $\pm$  0.0207 \\
  1.902 & 0.224 & 0.107 &  339.2 & -0.1958   $\pm$ 0.0607   $\pm$  0.0163 \\
         \hline \hline
      \end{tabular}
      \caption{ The coherent $A_{LU}$ in $Q^2$ bins}
      \label{table:Coh_Q2_BSA}
   \end{center}
\end{table}

### 2016-09-08 analysis-note-v2/RTPC.tex
Clarification on calibration and clean up

The variations with time of $TDC_{max/2}$ are shown in figure 
\ref{fig:Drift_run_number}, in which each point is a run of about two hours of 
data taking. One can see a non-negligible variation over the three months of 
data taking due to the changes in the experimental conditions. The main cause 
of these variations is probably the variation of proportions in the drift gas 
mixture which was not perfectly under control due to leaks in the thin parts of 
the detector.

@@

These functions are implemented in our reconstruction codes, and, as a result 
of adapting the drift speed with these functions, more good tracks and elastic 
events were identified comparing to the previous calibration set. Figure 
\ref{fig:Drift_run_number} also shows the gain percentages of the good tracks 
(GT) and the elastic events (El) in few runs. One can see a gradual improvement 
with time because of giving the right drift speed, which together with the 
drift paths gives more precise reconstructed hits chain. Therefore, the 
parameters of the reconstructed tracks ($sdist$, $edist$ ... etc) are more likely to pass the 
selection requirements and so more tracks are found to be good.

@@

\subsubsection{Drift paths parametrization}
A drift path is the trajectory that an electron follows after being 
released in the drift region. The standard software to calculate these paths is 
the MAGBOLTZ program \cite{MAGBOLTZ}. This program requires precise knowledge 
of the experimental conditions, such as the detector's geometry, the exact 
composition of the drift gas and the applied electric and magnetic fields. Thus 
any variation in these conditions would give inaccurate calculated drift paths.  
In the EG6 experiment, MAGBOLTZ has been used to calculate a first set of drift 
paths, but it was observed for the drift speed that we do not have a perfect 
knowledge of the RTPC's conditions. Therefore, the first set of the drift paths 
is not accurate. For this reason, we chose to calculate the drift paths in a 
way that does not require an exact knowledge of the conditions in the chamber 
and is instead based on our experimental data.

In our alternative method, we use the identified elastic ($e ~ ^{4}He 
\rightarrow e ~ ^{4}He$) events to extract the drift paths. For these events, 
the kinematics of the recoil $^{4}He$ are calculated from the scattered 
electron. Then the electrons' drift paths can be extracted from comparing the 
experimentally measured hits in the RTPC to the trajectory of a similar 
simulated $^{4}He$. In this technique, the drift paths are obtained 
independently of our knowledge of the exact conditions in the RTPC, in 
particular the electric field and the gas mixture have no impact.

Due to the magnetic field, the drift paths are not linear. This is handled by 
the following procedure, in which we perform the drift paths' extraction in 
two passes. In the first pass, we assume a linear correlation between the 
radius of emission $R$ (we work here in radial coordinates: $R$ and $\phi$) and the 
drift time, to link the GEANT4 hits to the measured hits. Through this association, we obtain an 
initial set of drift paths in the form of a $\Delta \phi$ distribution as function of time. We define $\Delta \phi$ the 
difference between the $\phi$ of the simulated hits and the $\phi$ of the pad 
which measured a hit. In the second pass, we refine the initial 
correlation between the radius and the time using the initial drift paths, and 
we extract final drift paths. To take into account the variations of the magnetic 
field and therefore the varying Lorentz angle, the extraction 
is made in $z$ bins along the RTPC.

The procedure steps are:
\begin{itemize}
\item We first select a sample of well identified elastic events in our experimental data. Using the
   kinematics of the scattered electrons in these events, we generate in GEANT4
   a sample of events identical to the measured ones. 
\item  For each couple of events (the experimental one and its simulated equivalent)
   we make a linear correlation between $R$, from simulation, and TDC from 
   real data, to link the GEANT4 hits to the measured ones. In our TPC, the 
   maximum $R$ equals 60 mm (close to the anode) at the trigger time, 
   (TDC$_{min}$ = 15 TDCs), and the minimum R equals 30 mm (close to the 
   cathode) at maximum TDC ($TDC_{max}$). Thus, the linear correlations can be 
   written as:
\begin{equation}
   R(TDC) = \frac{60-30}{15-TDC_{max}} (TDC-15) + 60.
\label{equ:R_TDC}
\end{equation}
We apply 3 TDCs-wide windows around the linear R-TDC correlation to associate 
the GEANT4 simulated hits to measured hits. Figure \ref{fig:R_correlation} 
shows the distribution of $R$ versus TDC for the selected simulated hits in a 
single $z$ bin.

\item From these selected hits, we construct the drift paths $\Delta \phi$ (= 
   $\phi_{sim.}$ - $\phi_{hit\_pad}$) versus TDC. The results can be seen in 
   figure \ref{fig:1st_pass_delta_phi}. $\Delta \phi$ at the anode (TDC = 15) 
   is not equal to zero because there is a drift in $\phi$ between the anode 
   (the first GEM layer at radial distance equal to 60 mm) and the readout pads 
   (at radial distance equal to 69mm).

\begin{figure}[tpb]
\centering
\includegraphics[scale=0.45]{fig_rtpc/updates/TdcR_check_p1_10.png}
\caption{(First pass). Distribution of $R$ as a function of TDC for the 
simulated hits located in one $z$ bin. The width of the bin is 10 mm, with the 
center at $z$ = 5 mm.}
\label{fig:R_correlation}
\end{figure}

\begin{figure}[tpb]
\centering
\includegraphics[scale=0.45]{fig_rtpc/TdcPhi_p1_10.png}
\caption{(First pass). Distribution of $\Delta \phi$ as a function of TDC for 
   in one $z$ bin. The width of the bin is 10 mm, 
with the center at $z$ = 5 mm. The black line represents a fit for 
$\Delta\phi$.  }
\label{fig:1st_pass_delta_phi}
\end{figure}

\begin{figure}[tbp]
\centering
\includegraphics[scale=0.45]{fig_rtpc/TdcR_p2_10.png}
\caption{(Second pass). The calculated R as a function of TDC is corrected by 
the $\Delta \phi$ relation extracted from the first pass.}
\label{fig:R_TDC}
\end{figure}

\item The correlation between R and TDC is refined using the extracted initial 
   $\Delta \phi$ as a function of TDC as:
\begin{equation}
   R (TDC) = R_{min} + \bigg[\sum\limits_{i=TDC_{Max}}^{TDC} \sqrt{DS^{2} - 
      R^{2}(i) \cdot \left(\frac{\partial \Delta \phi}{\partial TDC}(i) 
\right)^{2}}~~~~\bigg] \cdot TDC,
 \end{equation}
where DS is the average drift speed, equal to 0.7 mm/TDC ( 6.14 $\mu$m/ns), 
$R_{min}$ equals to 30 mm, $R(i)$ is the linear correlation defined in equation 
\ref{equ:R_TDC} and $\Delta \phi (i)$ is the fit in figure 
\ref{fig:1st_pass_delta_phi}. The sum is multiplied by a TDC (=114 ns) unit.  
The calculated R, figure \ref{fig:R_TDC}, is very close to a straight line, 
indicating that only two itterations are probably enough.

\item In the second pass, we use the newly found R(TDC) relation to construct new $\Delta \phi$ distributions, see figure \ref{fig:final_drift_paths}.

\begin{figure}[tp]
\centering
\includegraphics[scale=0.45]{fig_rtpc/FitResult_p2_10.png}
\caption{(Second pass). Distribution of $\Delta \phi$ as a function of TDC 
located in the same $z$ bin shown in figure \ref{fig:1st_pass_delta_phi}. The 
black line represents the final drift paths.}
\label{fig:final_drift_paths}
\end{figure}
\end{itemize}

We performed a third pass in which the R(TDC) correlation is refined using the 
drift paths from the second pass and new drift paths were extracted. As a 
result, we observed no difference between the drift paths of the second and the 
third passes. In other words, extracting the drift paths from two iterations 
gives us stable drift paths.

The first drift paths were extracted using the elastic events from the early 
1.204 GeV dataset. In order to ensure the consistency of these drift paths over 
the experimental running period, we extracted another set of drift paths using 
the elastic events of the 1.269 GeV runs. We observed no difference between the 
two sets of the drift paths, ensuring their stability over the experimental 
period and leaving the dependence on the TDC and $z$. The final drift paths are 
extracted using both datasets, 1.204 and 1.269 GeV. They take the form:
\vspace{-0.1in}
\begin{equation}
\Delta \phi (TDC, z)=  \sum\limits_{i=0}^{4} p_{i}(z)*TDC^{i},
\end{equation}
where the parameters, $p_{0}$, $p_{1}$, $p_{2}$, $p_{3}$ and $p_{4}$ are 
functions of $z$, as can be seen in Appendix \ref{app:RTPC_appendix}, table 
\ref{table:drift_paths}. The final drift paths are implemented in our 
reconstruction codes such that the reconstructed position of each hit becomes:
\begin{equation}
\phi_{exp}(z, TDC) = \phi _{hit\_pad} - \Delta \phi (z, TDC).
\end{equation}

@@

\subsubsection{summary}
The drift paths and drift speed were first extracted using the
MAGBOLTZ Monte Carlo simulation. One can find the detailed procedure of
a similar extraction in the BoNuS analysis note \cite{GEM_holes}.
Then, we improved this first result with many iterations after each change that 
would affect the drift paths, such as the beam-offset and gain calibrations of 
the RTPC. Moreover after each pass, the number of elastic events used for 
calibration increased, justifying by itself a new calibration. Overall, this 
was a long process that evolved and improved over the years and we concentrated 
in the note on the description of the final iteration of this calibration.

>>>>

The variations with time of $TDC_{max/2}$ are shown in figure 
\ref{fig:Drift_run_number}, in which each point is a run of about two hours of 
data taking. One can see a non-negligible variation over the three months of 
data taking due to the changes in the experimental conditions. The main cause 
of these variations is probably the variation of proportions in the drift gas 
mixture which was not perfectly under control due to leaks in the parts of 
the detector where the gas is contained by very thin aluminized mylar.

@@

The implementation of these parametrizations of the drift speed in our 
reconstruction codes resulted in more good tracks and elastic 
events identified comparing to the previous calibration sets. Figure 
\ref{fig:Drift_run_number} also shows these gain percentages of good tracks 
(GT) and elastic events (El) for a few runs. The gradual improvement 
with time is due to the improvement on the parameters of the reconstructed 
tracks ($sdist$, $edist$ ... etc), which are then more likely to pass the 
selection requirements.

@@

\subsubsection{Drift paths parametrization}

A drift path is the trajectory that an electron follows after being 
released in the drift region. The standard software to calculate these paths is 
the MAGBOLTZ program \cite{MAGBOLTZ}. This program requires precise knowledge 
of the experimental conditions, such as the detector's geometry, the exact 
composition of the drift gas and the applied electric and magnetic fields. Thus 
any variation in these conditions would give inaccurate calculated drift paths.  
In the EG6 experiment, MAGBOLTZ has been used to calculate a first set of drift 
paths, but it was found rapidly that our lack of knowledge on some of the RTPC's 
conditions was limiting the precision of the calculation. For this reason, we 
chose to further calibrate the drift paths with a method 
that does not require an exact knowledge of the conditions in the chamber 
and is instead based on the experimental data.

In this alternative method, we use the identified elastic ($e ~ ^{4}He 
\rightarrow e ~ ^{4}He$) events to extract the drift paths. For these events, 
the kinematic of the recoil $^{4}He$ is calculated from the scattered 
electron. Then the electrons' drift paths is extracted by comparing event
by event the experimentally measured hits in the RTPC to the 
trajectory of the corresponding simulated $^{4}He$. With this technique, the 
drift paths are obtained with limited impact from our knowledge of the 
conditions in the RTPC, in particular the method is completely independent 
from both electric field and gas mixture.

We perform the drift paths' extraction in two passes. 
In the first pass, we assume a linear correlation between the 
radius of emission $R$ (we work here in radial coordinates: $R$ and $\phi$) and the 
drift time to link GEANT4 hits along the trajectory to actually measured 
hits. Based on this association, we obtain a $\Delta \phi$ distribution 
as function of time. $\Delta \phi$ is the 
difference between the $\phi$ observed in GEANT4 and the $\phi$ of the pad 
which actually measured a correlated hit. Assuming we have an axial 
symmetry of the drift path, $\Delta \phi$ as a function of time indicates 
how much the electron drifted because of the magnetic field for any hit. 
This is therefore the relevant information to reconstruct the origin
of the hits measured in the detector.
However, the actual correlation between time and radius is not exactly
linear. So we perform a second pass, in which we refine the initial 
correlation between the radius and the time using the drift paths obtained 
in the first path. We then repeat the procedure of the first pass to extract 
the final drift paths. Because of the variations in the magnetic 
field along the detector, the extraction is made in $z$ bins.

The detailed steps of the procedure are as follow:
\begin{itemize}
\item We first select a sample of well identified elastic events in our 
   experimental data. Using the
   kinematics of the scattered electrons in these events, we generate in GEANT4
   a sample of events identical to the measured ones. 
\item  For each couple of events (the experimental one and its simulated equivalent)
   we make a linear correlation between $R$, from simulation, and TDC from 
   real data, to link the GEANT4 hits to the measured ones. In our TPC, the 
   maximum $R$ equals 60 mm (close to the anode) at the trigger time, 
   (TDC$_{min}$ = 15 TDCs), and the minimum $R$ equals 30 mm (close to the 
   cathode) at maximum TDC ($TDC_{max}$). Thus, the linear correlations can be 
   written as:
\begin{equation}
   R(TDC) = \frac{60-30}{15-TDC_{max}} (TDC-15) + 60.
\label{equ:R_TDC}
\end{equation}
We apply 3 TDCs-wide windows around the linear R-TDC correlation to associate 
the GEANT4 simulated hits to measured hits. Figure \ref{fig:R_correlation} 
shows the distribution of $R$ versus TDC for to select the matched hits.
\item From these selected matched hits, we obtain $\Delta \phi$ (= 
   $\phi_{sim.}$ - $\phi_{hit\_pad}$) as a function of time. The results can be seen in 
   figure \ref{fig:1st_pass_delta_phi}\footnote{$\Delta \phi$ at the anode (TDC = 15) 
   is not equal to zero because there is a drift in $\phi$ between the anode 
   (the first GEM layer at radial distance equal to 60 mm) and the readout pads 
   (at radial distance equal to 69mm).}. This figure is fitted to obtain the 
   parameterization of the drift path.
\begin{figure}[tpb]
\centering
\includegraphics[scale=0.45]{fig_rtpc/updates/TdcR_check_p1_10.png}
\caption{(First pass). Distribution of $R$ as a function of time (in TDC) for the 
simulated hits located in one $z$ bin. The width of the bin is 10 mm, with the 
center at $z$ = 5 mm.}
\label{fig:R_correlation}
\end{figure}
\begin{figure}[tpb]
\centering
\includegraphics[scale=0.45]{fig_rtpc/TdcPhi_p1_10.png}
\caption{(First pass). Distribution of $\Delta \phi$ as a function of time (in TDC) for 
   in one $z$ bin. The width of the bin is 10 mm, 
with the center at $z$ = 5 mm. The black line represents a fit for 
$\Delta\phi$.  }
\label{fig:1st_pass_delta_phi}
\end{figure} 
\begin{figure}[tbp]
\centering
\includegraphics[scale=0.45]{fig_rtpc/TdcR_p2_10.png}
\caption{(Second pass). The calculated $R$ as a function of time (in TDC) is corrected by 
the $\Delta \phi$ relation extracted from the first pass.}
\label{fig:R_TDC}
\end{figure}
\item The correlation between $R$ and the time is then refined using the extracted  
   $\Delta \phi$ as a function of time as:
\begin{equation}
   R (TDC) = R_{min} + \bigg[\sum\limits_{i=TDC_{Max}}^{TDC} \sqrt{DS^{2} - 
      R^{2}(i) \cdot \left(\frac{\partial \Delta \phi}{\partial TDC}(i) 
\right)^{2}}~~~~\bigg] \cdot TDC,
 \end{equation}
where DS is the average drift speed, equal to 0.7 mm/TDC ( 6.14 $\mu$m/ns), 
$R_{min}$ equals to 30 mm, $R(i)$ is the linear correlation defined in equation 
\ref{equ:R_TDC} and $\Delta \phi (i)$ is the fit in figure 
\ref{fig:1st_pass_delta_phi}. The sum is multiplied by a TDC unit (=114 ns).  
The calculated R, figure \ref{fig:R_TDC}, is almost linear, 
indicating that only two itterations are probably enough.
\item In the second pass, we use the newly found $R$(TDC) relation to construct 
new $\Delta \phi$ distributions, see figure \ref{fig:final_drift_paths} for our final
result.
\begin{figure}[tp]
\centering
\includegraphics[scale=0.45]{fig_rtpc/FitResult_p2_10.png}
\caption{(Second pass). Distribution of $\Delta \phi$ as a function of time (in TDC) 
located in the same $z$ bin shown in figure \ref{fig:1st_pass_delta_phi}. The 
black line represents the final drift paths.}
\label{fig:final_drift_paths}
\end{figure}
\end{itemize}

We performed a third pass in which the $R$(TDC) correlation is refined using the 
drift paths from the second pass. As a 
result, we observed no difference between the drift paths of the second and the 
third passes. In other words, extracting the drift paths from two iterations 
gives us stable results.

The first drift paths were extracted using the elastic events from the early 
1.204~GeV dataset. In order to ensure the consistency of these drift paths over 
the experimental running period, we extracted another set of drift paths using 
the elastic events of the 1.269 GeV dataset. We observed no difference between the 
two sets of the drift paths, ensuring the stability over the experimental 
period. The final drift paths are 
extracted using both datasets, 1.204 and 1.269 GeV. They take the form:
\vspace{-0.1in}
\begin{equation}
\Delta \phi (TDC, z)=  \sum\limits_{i=0}^{4} p_{i}(z)*TDC^{i},
\end{equation}
where the parameters, $p_{0}$, $p_{1}$, $p_{2}$, $p_{3}$ and $p_{4}$ are 
functions of $z$, as can be seen in Appendix \ref{app:RTPC_appendix}, table 
\ref{table:drift_paths}. The final drift paths are implemented in our 
reconstruction codes such that the reconstructed position of each hit becomes:
\begin{equation}
\phi_{exp}(z, TDC) = \phi _{hit\_pad} - \Delta \phi (z, TDC).
\end{equation}

@@

\subsubsection{Summary}
The drift paths and drift speed were first extracted using the
MAGBOLTZ Monte Carlo simulation. One can find the detailed procedure of
a similar extraction in the BoNuS analysis note \cite{GEM_holes}.
Then, we improved this first result with many iterations after each change that 
would affect the drift paths, such as the beam-offset and gain calibrations of 
the RTPC. Moreover after each pass, the number of elastic events used for 
calibration increased, justifying by itself a new calibration. Overall, this 
was a long process that evolved and improved over the years and we concentrated 
in the note on the description of the final iteration of this calibration.
